\chapter{Comparing NoSQL and Relational databases} 
% ! Chapter title doesn't fit the page.


% ? This section "shouldn't be too long, but enough to convince them."
% ! I'm targeting 1000 - 1250 words.
% ? This is NOT comparing *MongoDB* to relational databases.
% ? The brief mentions that it should be "generic just in case after seeing 
% ? your MongoDB database they decide to go with another software provider."

Industrial needs for fast and efficient data storage continue to grow exponentially year by year, and some of these needs fail to be met 
by typical relational database solutions. \textcite{corbelliniPersistingBigdataNoSQL2017} state that the types of data that require storage 
have changed since the first relational database systems of the 1970s, and as such, these more mature systems do not meet the requirements of
today's businesses with web-oriented systems storing magnitudes of unstructured data. The same can also be said of Internet of Things 
(IoT) systems, with  \textcite{gubbiInternetThingsIoT2013} stating that 'For the realization of a complete IoT vision, an efficient, secure,
scalable and market oriented computing and storage resourcing is essential.'

\para These efficient, secure and scalable resources exist in the form of NoSQL databases. Chapter 1 provided an overview of the various 
types of NoSQL databases, with this chapter instead offering direct comparisons between modern NoSQL solutions to more traditional relational 
database management systems in terms of their scalability and transactional models.

\section{The CAP theorem}
% TODO: Talk about the CAP theorem:
% ?     Consistency: Once data is written, it is available to all users of the system immediately.
% ?     Availability: The service is uninterrupted and without degradation for the majority of the time.
% ?     Partition Tolerance: Operations can be completed even if part of the network fails.
% * It's said that a distributed system can only ever achieve two of these.
% * https://www.ibm.com/think/topics/cap-theorem

\section{Scalability}
Both \textcite{ganiSurveyIndexingTechniques2016} and \textcite{katalBigDataIssues2013} state that the amount of data stored by companies 
is increasing at an extreme rate, forecasting that companies will likely be storing \textit{zettabytes} of data in the near future.
Because of this, \textcite{gubbiInternetThingsIoT2013}'s previously mentioned efficient and scalable storage solutions are needed to meet 
these ever-increasing demands.

\para Databases can be scaled both vertically and horizontally, with relational databases typically favouring vertical scaling and NoSQL 
databases favouring horizontal scaling, though both are possible in either database type. 

\subsection{Vertical scaling}
Scaling vertically means to upgrade the host device(s) of the database, which refers to the physical swapping or installation of 
parts such as CPUs, RAM and storage drives to accelerate data processing \autocite{cattellScalableSQLNoSQL2011}. 
A significant benefit of a vertically scaled database is that data will always be consistent as it is stored on a single device rather than
distributed across many over a network connection where data could be corrupted, lost or intercepted. This allows for total ACID compliance which is described further in Section \ref{subsec:ACID}. 

\para Despite this key benefit, vertical scaling has some downsides. Server components cannot be swapped while the system is online, meaning 
that scaling vertically \textbf{requires} partial or total system downtime, therefore decreasing availability. Furthermore, vertical scaling 
introduces a single point of failure to a system, where the powerful server hosting the database could experience any issue such as a power cut 
that would cause a total system outage. Continuous vertical scaling can also become very costly and bear diminishing returns, with the 
significant cost of top-grade server parts potentially outweighing the performance benefit of their installation.  

% ! A very text-heavy section (and CHAPTER, at that.). Add images maybe?

\subsection{Horizontal scaling}

% TODO: Talk about horizontal scalability.
% ?     Nodes and sharding

Scaling horizontally means to add more devices (nodes) to a network, making the database a \textbf{distributed database}. Rather 
than having an entire DBMS stored on a single device and single point of failure, horizontal scaling divides the processing and storage 
loads of the database over many nodes. 




\section{Transactional models}
Relational and NoSQL databases are governed by dichotomous transactional models that fundamentally change their appropriate use cases.
Relational databases are designed to adhere to the ACID model that prioritises absolute integrity and consistency at the expense of 
both availability and scalability, whereas most NoSQL databases adhere to the opposing BASE model, prioritising availability and scalability
over consistency. However, ACID is not exclusive to relational databases, and recent updates to certain NoSQL databases like MongoDB have 
allowed partial ACID compliance in combination with the scalability benefits of NoSQL.



\subsection{ACID}\label{subsec:ACID}
ACID is an acronym to describe the following key attributes:

% TODO: Talk about ACID:
% ?     Atomicity - A statement is fully executed or not executed at all.
% ?     Consistency - Transactions only occur in predefined ways. (Helps prevent impacts of corruption)
% ?     Isolation - Concurrent transactions don't interfere with each other.
% ?     Durability - Changes are saved, even in the event of system failure. 
% *    https://www.databricks.com/glossary/acid-transactions

\begin{longtable}{ | p{0.2\textwidth} | p{0.7\textwidth} | }
    \hline
    \cellcolor{blue!25}Property & \cellcolor{blue!25}Description\\
    \hline
    Atomicity & All transactions must be fully completed or not completed at all. If a transaction fails,
    the database must be reverted to its prior state with no changes made to the data.  \\
    \hline
    Consistency & Data must meet predefined integrity constraints, and remain consistent for all users even in the event 
    of simultaneous transactions.  \\
    \hline 
    Isolation & Transactions are executed sequentially, and do not interfere with each other even if they are simultaneous. \\
    \hline 
    Durability & Even in the event of system failure, the database must maintain all committed records. This means that 
    even if an error occurs, the results of any and all previous transactions must not be lost. \\
    \hline
    \caption{The key properties of ACID compliance. \autocite{awsACIDVsBASE,neo4jDataConsistencyModels2023}}\label{tab:ACID}
\end{longtable}

\noindent ACID is a very strict model that actively enforces a safe environment for data operations. Though, in maintaining 
such high security, a performance overhead exists with every transaction. This can massively impact the scalability of relational 
databases with complex join operations, which will grow slower the more users they have. 

\para ACID compliance is not exclusive to relational databases, however; in recent years, some NoSQL databases have incorporated ACID 
properties to gain the best of both worlds - high speed and scalability from being non-relational, and high security and integrity from 
ACID compliance. However, NoSQL databases cannot be fully ACID compliant due to their architecture, as horizontal scaling while maintaining 
 



\subsection{BASE}
% TODO: Talk about BASE:
% * A base is the opposite of an acid, hence the name.
% * Seems to be based on the *perception* of availability rather than the literal availability.
% ?     Basically Available - One transaction doesn't have to wait for another to complete first.
% ?     Soft State - The transitional state of records during concurrent updates. Finalised after all transactions complete.
% ?     Eventually Consistent - When all concurrent transactions are done, it will EVENTUALLY be consistent across all user perspectives.
% * https://aws.amazon.com/compare/the-difference-between-acid-and-base-database/

% ! I don't understand transient states and convergence; how can two opposite transactions merge to one eventual state?

BASE-compliant databases operate very differently to their ACID counterparts. Rather than the strict integrity and consistency procedures
expected by ACID, the properties of BASE databases are much more relaxed:

\begin{longtable}{ | p{0.3\textwidth} | p{0.6\textwidth} | }
    \hline
    \cellcolor{blue!25}Property & \cellcolor{blue!25}Description\\
    \hline
    Basically Available & Systems will continue to operate even if failures occur due to redundancy and replication.  \\
    \hline
    Soft State & Replicas do not have to be entirely consistent with each other. This means the system 
    can be in two different states simultaneously, which is known as a transient state.\\
    \hline 
    Eventually Consistent & While updates are not occurring, transient states will converge. \\
    \hline
    \caption{BASE properties. \autocite{corbelliniPersistingBigdataNoSQL2017,cattellScalableSQLNoSQL2011,neo4jDataConsistencyModels2023,awsACIDVsBASE}}\label{tab:BASE}
\end{longtable}

\noindent By forfeiting the strict regulations and associated performance overheads of ACID compliance, BASE-compliant databases can 
theoretically operate at much higher speeds by sacrificing their immediate consistency.


% ! While you've definitely described all the terms well, have you actually even compared the DB types at all?

