\chapter{Database design and implementation}

% MongoDB, an acronym from "hu\textbf{mongo}us \textbf{DB}", aims to address some of SQL's antiquation issues\dots

\section{Target problem}
\subsection{Clientele}
IoThings are a UK-based small-to-medium sized enterprise specialising in Internet of Things home automation systems.
Specifically, they install sensors and use MQ Telemetry Transport (MQTT) to collect their activation data and store it in 
a MongoDB database for analytical purposes. These sensors can be used for a variety of reasons, such as:

\begin{itemize}
    \item Temperature control
    \item Lighting control 
    \item Entertainment management 
    \item Home security
\end{itemize}

\subsection{Existing systems}
IoThings already have an extensive data storage solution for their Enterprise Resource Planning (ERP), Customer Relationship 
Management (CRM), finances, order processing / sales, and logistics. However, they do not presently have an optimal solution for 
their MQTT sensor activation data.  


\subsection{Requested solution}
Therefore, IoThings wish to extend their current data storage implementation with an additional NoSQL database to contain their 
sensor activation data for the eventual purpose of giving feedback to their users about device usage. In addition to the 
creation of this database, the client also requests the development of an API and frontend for accessing and processing the 
stored data.

\para Due to data security and privacy concerns, it is not possible for IoThings to provide data of their own. Therefore, 
the solution developed in this report will use synthetic data which aims to replicate the data which would theoretically 
be stored.